Symbolic linearization of equations of motion of constrained multibody systems

### 2014-01-20 PetersonGedeHubbard2012.tex
Add figure and fix math mode errors.
Addressed most of Mont's suggestions
About 1 page worth of corrections remains, as well as a minor change to the
figure and a few others.

\begin{abstract}
% Context
Many common dynamic mechanical systems are subject to configuration or velocity
constraints.
% Need  (what we have and what we want)
Analyses of such systems often require linearized forms of the motion
equations.
% Task
To address this issue, we developed a procedure for organizing the constraint
and motion equations and their subsequent linearization.  This procedure was
specifically developed for equations of motion generated by Kane's method, but
is compatible with any method as long as the following are true: only ordinary
differential equations are required to describe the system, and independent and
dependent states remain in the final equations (i.e. no DAE's, dependent
variables are not substituted away).
% Object of the document
Following a brief review of Kane's method and the structure of the equations it
generates, we present the procedure to symbolically linearize the nonlinear
motion equations of constrained multibody systems, and illustrate it with an
example of the rolling disk.

@@

\section{Introduction}
\label{sec:intro}
Design of linear state feedback controllers and first order stability analysis
of equilibria are two of the most common uses of multibody system dynamics
models. Both of these uses require linearized equations of motion. For
unconstrained systems, obtaining the linearized equations of motion is as
simple as arranging the equations of motion in first order form ($\dot{x} =
f(x, t)$) and evaluating the Jacobian of the right hand side function, with
respect to the states, evaluated at the equilibria point of interest.  However,
when the system in questions has constraints, i.e., not all the states are
independent, this process is not as straightforward.

For a system with constraints, Kane's dynamical equations\cite{Kane1985} are
first order ODE's (in the generalized speeds) which include both the
independent and dependent states (assuming the dependent states have not been
algebraically eliminated). The kinematical and dynamical differential equations
can be solved for the time derivatives of the generalized coordinates and
generalized speeds, respectively. Taken together, it is tempting to think that
this system of first order differential equations could be linearized in the
same fashion as an unconstrained system.  Unfortunately, this is not the case.
Linearizing these equations in the same manner as one would for an
unconstrained system (by directly computing the Jacobian) is incorrect and will
yield incorrect linearized equations (and subsequently incorrect eigenvalues,
in the case of a stability analysis). We illustrate this by means of a familiar
example, the rolling disk, formulated with a non-traditional set of coordinates
and generalized speeds which are not all independent. We provide a systematic
method to correctly generate the linear equations of motion for constrained
systems which have been found using Kane's method (or any other method which
results in equations with the same form). We apply this method to the rolling
disk example to illustrate that it yields eigenvalues that match previously
published results.

We consider three levels of constraints: configuration, velocity, and
acceleration. Configuration (holonomic) constraints limit the location or
orientation of parts of the system, relative to the external world or other
parts of the system. Velocity (nonholonomic of time differentiated holonomic)
constraints limit the speeds at which the configuration can change.
Velocity constraints most often appear in systems where there is rolling
without slip or there are closed kinematic loops. For this procedure,
acceleration level constraints are assumed to be time differentiated velocity
constraints.

% TODO: Improve literature review.
Bottema appears to be the first author to have shown that special
considerations are needed for linearizing nonholonomic systems
\cite{Bottema1949}. While other authors have examined linearization of
nonholonomic systems, we have found that most techniques are not directly
applicable to equations formed using Kane's method, and while we do not doubt
that they \textit{could} be applied correctly, the details of how to do so, as
far as the authors are aware, have not been explicitly presented. Kang et al.
\cite{Kang2003} and Negrut and Ortiz \cite{Negrut2006} have both explored
linearization, but these methods have been developed for equations of motion
found using Lagrange's method, which contain Lagrange multipliers (which are
not present in equations of motion generated from Kane's method). Further, both
of these papers lack a discussion of basic concepts such as how many quantities
are independent, and in the case of eigenvalue computation, how many
eigenvalues should be present. Minaker and Rieveley \cite{Minaker2010} and
Schwab and Meijaard \cite{Schwab2003} both have developed techniques for
generating equations of motion (and linear forms thereof) using Lagrange's
method; the work of Schwab and Meijaard can be applied symbolically while
Minaker and Reiveley's approach is numeric in nature. Neither discuss the
choice of generalized speeds as in Kane's method: both form the equations of
motion in terms of second time derivatives of the coordinates.  Further,
neither of these works address constraints which cannot be eliminated (e.g.,
nonlinear configuration constraints), nor do they address what approach should
be taken when a single choice of dependent states may not suffice. While
Neuman presents a symbolic (as opposed to numeric) formulation for
linearization of the Q-matrix formulation of the Lagrangian\cite{Neuman1984},
his work makes no mention of constrained systems and attempts to contact the
author regarding the software (algebraic robot modeler -- Arm) as well as
internet searches for the software have been fruitless.

@@

% TODO:reiterate limitations of model
This paper presents linear, first order differential equations relating the
time derivatives of \textit{all} speeds and coordinates to the independent
speeds and coordinates, for any point of linearization.  A provided example
demonstrates the necessity of these equations.  We demonstrate how a reduced
portion of these linearized equations can be used to perform standard stability
analysis (i.e., for control system design or linear stability analysis).  The
formalism presented here is generic enough to cover most examples of
time-varying constrained multibody systems with arbitrary external inputs and
arbitrary specified quantities.
% TODO: electronic supplementary material with examples for other methods
% TODO: rewrite following sentences after e.s.m.
The procedure was created for systems whose equations of motion have been
derived with Kane's method. Although also applicable to dynamic system
equations formulated using other methods, it is restricted to systems which can
be completely described by a set of ODE's (i.e., DAE's needn't be solved).

We begin with a review of Kane's method for generating equations of motion is
reviewed first, in order to properly orient the reader to the format and some
of the qualities of the generated equations. Next, we apply Kane's method to
generate the equations of motion for the rolling disk, but we purposefully
introduce a dependent coordinate and three dependent speeds in the formulation.
The need for a different linearization technique is demonstrated with this
example. In response to this need, we present the newly developed linearization
procedure and its derivation. This new linearization procedure is applied the
same rolling disk example to illustrate its validity and an example of its
application. Finally, we discuss other details and nuances of the procedure and
its use, as well as directions for future work.

\section{Kane's Method, briefly}
\label{sec:kane_method}
To familiarize the reader with Kane's method, some concepts relating to its use
will be described.  The first is the use of generalized speeds in addition to
generalized coordinates.  The second concept is in how these generalized speeds
can be used to project permissible motions of the system, and how this removes
the need to consider non-contributing (internal constraint) forces.  Then, the
mathematical steps required to use Kane's method are shown and described.

When using Lagrange's method, generalized coordinates are used to describe the
configuration of a system while the time derivatives of the generalized
coordinates describe the velocities of a system.  Kane's method allows for the
velocities of a system to be written in terms of generalized speeds, which are
not required to be the time derivatives of coordinates (although such a
definition is permitted).

The benefit of choosing generalized speeds to be other than simply the
derivatives of the corresponding generalized coordinates can be seen in the
following example. If the orientation of a rigid body were to be described
using using Euler parameters or quaternions as coordinates, its angular
velocity in terms of time derivatives of those coordinates is relatively
complicated; it also requires a velocity constraint. Using generalized speeds
allows for the angular velocity to be expressed more simply as
\begin{align}
\label{eq:ex1}
{^N}\bar{\omega}^B = u_1 \hat{b}_1 + u_2 \hat{b}_2 + u_3 \hat{b}_3
\end{align}
or, the angular velocity of rigid body $B$ in the reference frame $N$ is the
defined as the sum of three generalized--speed basis--vector products.
The derivatives of these generalized speeds will appear in the equations
of motion, rather than the twice time differentiated generalized coordinates
($\ddot{\mathbf{q}}$'s).

@@

Use and understanding of Kane's method requires the use and understanding of
partial velocities and partial angular velocities. These are simply the
partial derivatives of the (angular) velocity vectors with respect to the
generalized speeds. For example, for the angular velocity in equation
(\ref{eq:ex1}), the partial angular velocities of body $B$ in reference frame
$N$ are
\[
{^N}\bar{\omega}^B_{1} = \hat{b}_1 \quad \quad
{^N}\bar{\omega}^B_{2} = \hat{b}_2 \quad \quad
{^N}\bar{\omega}^B_{3} = \hat{b}_3
\]
and it can be seen that the partial velocities represent the direction of
motion associated with each generalized speed. The geometric interpretations of
these partial (angular) velocities are discussed at length by
Lesser\cite{Lesser1992}.

Taking the dot product of each partial velocity and both sides of Newton's
second law, $\bar{F}=m\bar{a}$ (or the rotational equivalent), only terms that
are parallel to the partial velocities remain; one scalar equation is generated
for each generalized speed. An important byproduct is that internal constraint forces
(non-contributing forces) imposed by objects such as pin or sliding joints no
longer need to be considered. The reason for this is that partial velocities
are \textit{by construction} perpendicular to forces of constraint. Constraint
forces can be included in the formulation, but will not appear in the final
equations of motion because of this very desirable property.

@@

In order to properly apply Kane's method, the velocities of each component in
the system need to be written as linear combinations of the generalized speeds.
The translational velocity of the particles are
\begin{align}
\label{eq:particle_translational}
{^N}\bar{v}^{P_i} = \sum^o_{j=1} {^N}\bar{v}^{P_i}_{j} u_j + {^N}\bar{v}^{P_i}_t
\quad \quad i=1,...,g
\end{align}
while the translational velocity of rigid body mass centers are
\begin{align}
\label{eq:rb_translational}
{^N}\bar{v}^{B_i^*} = \sum_{j=1}^o {^N}\bar{v}^{B_i^*}_{j} u_j +
{^N}\bar{v}^{B_i^*}_t \quad \quad i=1,...,h
\end{align}
and finally the translational velocity of points which have applied forces are
\begin{align}
\label{eq:points_translational}
{^N}\bar{v}^{Q_i} = \sum_{j=1}^o {^N}\bar{v}^{Q_i}_{j} u_j + {^N}\bar{v}^{Q_i}_t
\quad \quad i=1,...,k
\end{align}
Next, the angular velocity of rigid bodies are
\begin{align}
\label{eq:rb_rotational_bodies}
{^N}\bar{\omega}^{B_i} = \sum_{j=1}^o {^N}\bar{\omega}^{B_i}_{j} u_j +
{^N}\bar{\omega}^{B_i}_t \quad \quad i=1,...,h
\end{align}
and the angular velocity of reference frames which have applied torques are
\begin{align}
\label{eq:rb_rotational_frames}
{^N}\bar{\omega}^{E_i} = \sum_{j=1}^o {^N}\bar{\omega}^{E_i}_{j} u_j +
{^N}\bar{\omega}^{E_i}_t \quad \quad i=1,...,c
\end{align}

@@

For this system, we consider configuration constraints of the form
\begin{align}
\label{eq:configuration_constraints}
\mathbf{f}_c(\mathbf{q}, t) = \mathbf{0}
\end{align}
where the terms $\mathbf{f}_c$ and $\mathbf{0}$ have the shape $l \times 1$.
The velocity constraints are taken to have the form
\begin{align}
\label{eq:velocity_constraints}
\mathbf{f}_v(\mathbf{q}, \mathbf{u}, t) = \mathbf{0}
\end{align}
where the terms $\mathbf{f}_v$ and $\mathbf{0}$ have the shape $m \times 1$,
and $\mathbf{f}_v$ depends linearly on $\mathbf{u}$.  It is important that the
velocity constraints include the time-differentiated configuration constraints
or equivalent constraints which produce the same behavior. Furthermore,
velocity constraints should be written such that generalized speeds and
coordinates appear, but $\dot{\mathbf{q}}$ terms do not.  Finally, acceleration
constraints are typically time-differentiated velocity constraints (or
similarly formed from kinematic considerations) and are assumed to have the
form
\begin{align}
\label{eq:acceleration_constraints}
\mathbf{f}_a(\mathbf{q}, \dot{\mathbf{q}}, \mathbf{u}, \dot{\mathbf{u}}, t) =
\mathbf{0}
\end{align}
In contrast to the velocity constraints, our formulation permits the
acceleration constraints to involve $\dot{\mathbf{q}}$ terms (alternatively
they can be eliminated by appealing to the kinematic differential equations).

@@

All the terms in (\ref{eq:kanes_eq}) are now defined, but this equation is
valid only if there are no velocity constraints. If there are velocity
constraints, the equations must be transformed. As we present them, Kane's
nonholonomic dynamical equations will be of dimension $o - m$, where $o$ is the
total number of generalized speeds and $m$ is the number of velocity
constraints. However, these $o-m$ equations generally involve all $o$ speeds
(we assume the dependent speeds and their time derivatives have \textit{not}
been algebraically eliminated). In order to obtain the nonholonomic dynamical
equations, however, a choice must be made as to which speeds will be dependent.
The methods for doing this are outside the scope of this paper, but it is
important that whatever selection is made avoids singularities which can occur
in certain configurations or sets of system parameters. For a careful treatment
of this, we encourage the reader to examine the work of Reckdahl
\cite{Reckdahl1996}.

We assume the $o$ generalized speeds are ordered $\mathbf{u} = [\mathbf{u}_i,
\mathbf{u}_d]^T$, where $\mathbf{u}_i\in\mathbf{R}^{o-m}$ and
$\mathbf{u}_d\in\mathbf{R}^{m}$ are the independent and dependent speeds,
respectively. The velocity constraint Jacobian matrix $B$ can be similarly
rearranged as $\mathbf{B} = [\mathbf{B}_i, \mathbf{B}_d]$, where
$\mathbf{B}_i\in \mathbf{R}^{m \times o-m}$ and $\mathbf{B}_d \in \mathbf{R}^{m
\times m}$. With this in mind, equation (\ref{eq:constraint_Bu0}) can be
written as
\begin{align}
\mathbf{B}_i \mathbf{u}_i + \mathbf{B}_d \mathbf{u}_d + \mathbf{f}_{vt}(\mathbf{q}, t) &= \mathbf{0} \\
\implies \mathbf{u}_d &= -\mathbf{B}_d^{-1} \mathbf{B}_i \mathbf{u}_i - \mathbf{f}_{vt}(\mathbf{q}, t)
\end{align}
For convenience, we define
\begin{align}
\label{eq:constraint_A}
\mathbf{A} \triangleq - \mathbf{B}_{d}^{-1} \mathbf{B}_i
\end{align}
and note that $\mathbf{A}\in\mathbf{R}^{m \times (o - m)}$. To constrain the
generalized active and inertia forces, we partitioned them in a similar
manner
\begin{align}
\label{eq:F_Fstar_rewrite}
\mathbf{F} = \begin{bmatrix} \mathbf{F}_i \\ \mathbf{F}_d \end{bmatrix} \quad
\quad \mathbf{F}^* = \begin{bmatrix} \mathbf{F}^*_i \\ \mathbf{F}^*_d
\end{bmatrix}
\end{align}
allowing the nonholonomic generalized forces to be written:
\begin{align}
\label{eq:F_Fstar_nonholonomic}
\tilde{\mathbf{F}} = \mathbf{F}_i + \mathbf{A}^T \mathbf{F}_d \quad \quad
\tilde{\mathbf{F}}^* = \mathbf{F}_i^* + \mathbf{A}^T \mathbf{F}_d^*
\end{align}
which allows Kane's nonholonomic dynamical equations to be written as
\begin{align}
\label{eq:kanes_eq_nonholonomic}
\tilde{\mathbf{F}} + \tilde{\mathbf{F}}^* = \mathbf{0}
\end{align}
This process of constraining the equations of motion reduces the number of
equations from $o$ to $o-m$. However, the (\ref{eq:kanes_eq_nonholonomic}) still
contain $o$ unknowns ($\dot{\mathbf{u}}_1,...,\dot{\mathbf{u}}_o$). By
augmenting (\ref{eq:kanes_eq_nonholonomic}) with the acceleration constraint
equations (\ref{eq:acceleration_constraints}), we obtain $o$ equations in
$o$ unknowns. Note that equations (\ref{eq:kanes_eq_nonholonomic}) and
(\ref{eq:acceleration_constraints}) are linear in the $\dot{u}_i$'s. The
augmented system of equations is
\begin{align}
\label{eq:F_Fstar_f_a}
\begin{bmatrix}
\tilde{\mathbf{F}} + \tilde{\mathbf{F}}^* \\
\mathbf{f}_a (\mathbf{q}, \dot{\mathbf{q}}, \mathbf{u}, \dot{\mathbf{u}}, t)
\end{bmatrix} = \mathbf{0}
\end{align}

@@

Not all of the $n$ generalized coordinates are independent due to the $l$
configuration level constraints (e.g., holonomic constraints resulting from
closed kinematic loops). Similarly, not all of the $o$ generalized speeds are
independent due to the $m$ velocity level constraints.  Velocity level
constraints result from time differentiated configuration constraints or
nonholonomic constraints (e.g., no-slip rolling).  The $m$ acceleration level
constraints are assumed to be obtained by time differentiating the $m$ velocity
level constraints.  Just as Kane's method does not require $n = o$,
neither does our linearization procedure. The $s$ exogenous inputs include
control system inputs and/or external disturbance forces or torques applied to
the system.

The kinematic differential equations are organized into the quantities
$\mathbf{f}_0$ and $\mathbf{f}_1$; terms depending upon $\dot{\mathbf{q}}$ are
placed into $\mathbf{f}_0$, all other terms are placed into $\mathbf{f}_1$.
Similarly, the dynamic differential equations are organized into the
quantities $\mathbf{f}_2$ and $\mathbf{f}_3$; terms depending upon
$\dot{\mathbf{u}}$ are placed into $\mathbf{f}_2$, all other terms are placed
into $\mathbf{f}_3$. The motivation for this organizational scheme will be made
apparent in Section \ref{sec:derivations}.

>>>>

\begin{abstract}
% Context
Many common dynamic mechanical systems are subject to configuration or velocity
constraints.
% Need  (what we have and what we want)
Analyses of such systems often require linearized forms of the motion
equations.
% Task
To address this issue, we developed a procedure for organizing the constraint
and motion equations and their subsequent linearization.  This procedure was
developed specifically for equations of motion generated by Kane's method, but
is compatible with any method as long as only ordinary differential equations
are required to describe the system, and independent and dependent states
remain in the final equations (i.e. no DAE's, dependent variables are not
substituted away).
% Object of the document
Following a brief review of Kane's method and the structure of the equations it
generates, we present the procedure to symbolically linearize the nonlinear
motion equations of constrained multibody systems, and illustrate it with an
example of the rolling disk.

@@

\section{Introduction}
\label{sec:intro}
Design of linear state feedback controllers and first order stability analysis
of equilibria are two of the most common uses of multibody system dynamics
models. Both require linearized equations of motion. For unconstrained systems,
obtaining the linearized equations of motion is as simple as arranging the
equations of motion in first order form ($\dot{x} = f(x, t)$) and calculating
the Jacobian of the right hand side function, with respect to the states,
evaluated at the equilibria point of interest.  However, when the system in
questions has constraints, i.e., not all the states are independent, this
process is not as straightforward.

For a system with constraints, Kane's dynamical equations\cite{Kane1985} are
first order ODE's (in the generalized speeds) which include both the
independent and dependent states (assuming the dependent states have not been
algebraically eliminated). The kinematical and dynamical differential equations
can be solved for the time derivatives of the generalized coordinates and
generalized speeds, respectively. Taken together, it is tempting to think that
this system of first order differential equations could be linearized in the
same fashion as an unconstrained system.  Unfortunately, this is not the case.
Linearizing these equations in the same manner as one would for an
unconstrained system (by directly computing the Jacobian) is incorrect and will
yield incorrect linearized equations (and subsequently incorrect eigenvalues,
in the case of a stability analysis). We illustrate this by means of a familiar
example, the rolling disk, formulated with a non-traditional set of coordinates
and generalized speeds which are not all independent. We then provide a
systematic method to correctly generate the linear equations of motion for
constrained systems which have been found using Kane's method (or any other
method that results in equations with the same form). We apply this method to
the rolling disk example to illustrate that it yields eigenvalues that match
previously published results.

We consider three levels of constraints: configuration, velocity, and
acceleration. Configuration (holonomic) constraints limit the location or
orientation of parts of the system, relative to the external world or other
parts of the system. Velocity (nonholonomic or time differentiated holonomic)
constraints limit the speeds at which the configuration can change.
Velocity constraints often appear where there is rolling without slip or there
are closed kinematic loops. For this procedure, acceleration level constraints
are assumed to be time differentiated velocity constraints.

% TODO: Improve literature review.
Bottema appears to be the first author to have shown that special
considerations are needed for linearizing nonholonomic systems
\cite{Bottema1949}. While other authors have examined linearization of
nonholonomic systems, we have found that most techniques are not directly
applicable to equations formed using Kane's method. While we do not doubt that
they \textit{could} be applied correctly, the details of how to do so, as far
as the authors are aware, have not been explicitly presented. Kang et al.
\cite{Kang2003} and Negrut and Ortiz \cite{Negrut2006} have both explored
linearization, but their methods have been developed for equations of motion
found using Lagrange's method, which contain Lagrange multipliers ( not present
in equations of motion generated using Kane's method). Further, both of these
papers lack a discussion of basic concepts such as how many quantities are
independent, and in the case of eigenvalue computation, how many eigenvalues
should be present. Minaker and Rieveley \cite{Minaker2010} and Schwab and
Meijaard \cite{Schwab2003} have developed techniques for generating equations
of motion (and linear forms thereof) using Lagrange's method; the work of
Schwab and Meijaard can be applied symbolically while Minaker and Reiveley's
approach is numeric in nature. Neither discusses the choice of generalized
speeds as in Kane's method: both form the equations of motion in terms of
second time derivatives of the coordinates.  Further, neither of these works
addresses constraints which cannot be eliminated (e.g., nonlinear configuration
constraints), nor do they address what approach should be taken when a single
choice of dependent states may not suffice.  Neuman \cite{Neuman1984} presents
a symbolic (as opposed to numeric) formulation for linearization of the
Q-matrix formulation of the Lagrangian, but makes no mention of constrained
systems and attempts to contact the author regarding the software (algebraic
robot modeler -- Arm) as well as internet searches for the software have been
fruitless.

@@

% TODO:reiterate limitations of model
This paper presents linear, first order differential equations relating the
time derivatives of \textit{all} speeds and coordinates to the independent
speeds and coordinates, for any point of linearization.  A provided example
demonstrates the need for these equations.  We demonstrate how a reduced
set of these linearized equations can be used to perform standard stability
analysis (i.e., for control system design or linear stability analysis).  The
formalism presented here is generic enough to cover most examples of
time-varying constrained multibody systems with arbitrary external inputs and
arbitrary specified quantities.
% TODO: electronic supplementary material with examples for other methods
% TODO: rewrite following sentences after e.s.m.
The procedure was created for systems whose equations of motion have been
derived with Kane's method. Although also applicable to dynamic system
equations formulated using other methods, it is restricted to systems which can
be completely described by a set of ODE's (i.e., DAE's needn't be solved).

We begin with a review of Kane's method for generating equations of motion, in
order to properly orient the reader to the format and some of the qualities of
the generated equations. Next, we apply Kane's method to generate the equations
of motion for the rolling disk, but we purposefully introduce a dependent
coordinate and three dependent speeds in the formulation.  The need for a
different linearization technique is demonstrated with this example. In
response to this need, we present the newly developed linearization procedure
and its derivation. This new linearization procedure is applied to the same
rolling disk example to illustrate its validity and an example of its
application. Finally, we discuss other details and nuances of the procedure and
its use, as well as directions for future work.

\section{Kane's Method, briefly}
\label{sec:kane_method}
To familiarize the reader with Kane's method, several concepts relating to its
use are discussed: the use of generalized speeds in addition to generalized
coordinates; how these generalized speeds can be used to project permissible
motions of the system, and how this removes the need to consider
non-contributing (internal constraint) forces.  Then, the mathematical steps
required to use Kane's method are illustrated.

In Lagrange's method, generalized coordinates describe the configuration of a
system while time derivatives of the generalized coordinates characterize the
velocities of a system.  Kane's method allows velocities to be written in terms
of generalized speeds, which are not required to be the time derivatives of
coordinates (although such a definition is permitted).

The benefit of choosing generalized speeds other than simply derivatives of
corresponding generalized coordinates can be seen in the following example. If
the orientation of a rigid body were described using Euler parameters or
quaternions as coordinates, its angular velocity in terms of time derivatives
of those coordinates is relatively complicated; it also requires a velocity
constraint. Using generalized speeds allows for the angular velocity to be
expressed more simply as
\begin{align}
\label{eq:ex1}
{^N}\bar{\omega}^B = u_1 \hat{b}_1 + u_2 \hat{b}_2 + u_3 \hat{b}_3
\end{align}
or, the angular velocity of rigid body $B$ in the reference frame $N$ is
defined as the sum of three generalized--speed basis--vector products.
The derivatives of these generalized speeds will appear in the equations
of motion, rather than the twice time differentiated generalized coordinates
($\ddot{q}$'s).

@@

Use and understanding of Kane's method requires the use and understanding of
\textit{partial velocities} and \textit{partial angular velocities}. These are
simply the partial derivatives of the (angular) velocity vectors with respect
to the generalized speeds. For example, for the angular velocity in equation
(\ref{eq:ex1}), the three partial angular velocities of body $B$ in reference
frame $N$ are
\[
{^N}\bar{\omega}^B_{1} = \hat{b}_1 \quad \quad
{^N}\bar{\omega}^B_{2} = \hat{b}_2 \quad \quad
{^N}\bar{\omega}^B_{3} = \hat{b}_3
\]
and it can be seen that the partial velocities represent the direction of
motion associated with each generalized speed. The geometric interpretations of
these partial (angular) velocities are discussed at length by
Lesser\cite{Lesser1992}.

Taking the dot product of each partial velocity with both sides of Newton's
second law, $\bar{F}=m\bar{a}$ (or the rotational equivalent), only terms that
are parallel to the partial velocities remain; one scalar equation is generated
for each generalized speed. An important byproduct is that internal constraint forces
(non-contributing forces) imposed by objects such as pin or sliding joints no
longer need to be considered. The reason for this is that partial velocities
are \textit{by construction} perpendicular to forces of constraint. Constraint
forces can be included in the formulation, but will not appear in the final
equations of motion because of this very desirable property.

@@

In order to properly apply Kane's method, the velocities of each component in
the system need to be written as linear combinations of the generalized speeds.
The translational velocities of the particles are
\begin{align}
\label{eq:particle_translational}
{^N}\bar{v}^{P_i} = \sum^o_{j=1} {^N}\bar{v}^{P_i}_{j} u_j + {^N}\bar{v}^{P_i}_t
\quad \quad i=1,...,g
\end{align}
while the translational velocities of rigid body mass centers are
\begin{align}
\label{eq:rb_translational}
{^N}\bar{v}^{B_i^*} = \sum_{j=1}^o {^N}\bar{v}^{B_i^*}_{j} u_j +
{^N}\bar{v}^{B_i^*}_t \quad \quad i=1,...,h
\end{align}
and finally the translational velocities of points which have applied forces are
\begin{align}
\label{eq:points_translational}
{^N}\bar{v}^{Q_i} = \sum_{j=1}^o {^N}\bar{v}^{Q_i}_{j} u_j + {^N}\bar{v}^{Q_i}_t
\quad \quad i=1,...,k
\end{align}
Next, the angular velocities of rigid bodies are
\begin{align}
\label{eq:rb_rotational_bodies}
{^N}\bar{\omega}^{B_i} = \sum_{j=1}^o {^N}\bar{\omega}^{B_i}_{j} u_j +
{^N}\bar{\omega}^{B_i}_t \quad \quad i=1,...,h
\end{align}
and the angular velocities of reference frames where a torque is being applied
are
\begin{align}
\label{eq:rb_rotational_frames}
{^N}\bar{\omega}^{E_i} = \sum_{j=1}^o {^N}\bar{\omega}^{E_i}_{j} u_j +
{^N}\bar{\omega}^{E_i}_t \quad \quad i=1,...,c
\end{align}

@@

For this system, we consider configuration constraints of the form
\begin{align}
\label{eq:configuration_constraints}
\mathbf{f}_c(\mathbf{q}, t) = \mathbf{0}
\end{align}
where $\mathbf{f}_c$ and $\mathbf{0}$ have shape $l \times 1$.  The velocity
constraints have the form
\begin{align}
\label{eq:velocity_constraints}
\mathbf{f}_v(\mathbf{q}, \mathbf{u}, t) = \mathbf{0}
\end{align}
where $\mathbf{f}_v$ and $\mathbf{0}$ have shape $m \times 1$,
and $\mathbf{f}_v$ depends linearly on $\mathbf{u}$.  It is important that the
velocity constraints include the time-differentiated configuration constraints
or equivalent constraints which produce the same behavior. Furthermore,
velocity constraints should be written such that generalized speeds and
coordinates appear, but $\dot{\mathbf{q}}$ terms do not.  Finally, acceleration
constraints are typically time-differentiated velocity constraints (or
similarly formed from kinematic considerations) and are assumed to have the
form
\begin{align}
\label{eq:acceleration_constraints}
\mathbf{f}_a(\mathbf{q}, \dot{\mathbf{q}}, \mathbf{u}, \dot{\mathbf{u}}, t) =
\mathbf{0}
\end{align}
In contrast to the velocity constraints, our formulation permits the
acceleration constraints to involve $\dot{\mathbf{q}}$ terms (alternatively
they can be eliminated by appealing to the kinematic differential equations).

@@

All the terms in (\ref{eq:kanes_eq}) are now defined, but this equation is
valid only if there are no velocity constraints. If there are velocity
constraints, the equations must be transformed. As we present them, Kane's
nonholonomic dynamical equations will be of dimension $o - m$, where $o$ is the
total number of generalized speeds and $m$ is the number of velocity
constraints. However, these $o-m$ equations generally involve all $o$ speeds
(we assume the dependent speeds and their time derivatives have \textit{not}
been algebraically eliminated). To obtain the nonholonomic dynamical equations,
however, a choice must be made as to which speeds will be dependent.  The
methods for doing this are beyond the scope of this paper, but it is important
that whatever selection is made avoids singularities which can occur in certain
configurations or sets of system parameters. For a careful treatment , we
encourage the reader to examine the work of Reckdahl \cite{Reckdahl1996}.

We assume the $o$ generalized speeds are ordered $\mathbf{u} = [\mathbf{u}_i,
\mathbf{u}_d]^T$, where $\mathbf{u}_i\in\mathbf{R}^{o-m}$ and
$\mathbf{u}_d\in\mathbf{R}^{m}$ are the independent and dependent speeds,
respectively. The velocity constraint Jacobian matrix $B$ can be similarly
rearranged as $\mathbf{B} = [\mathbf{B}_i, \mathbf{B}_d]$, where
$\mathbf{B}_i\in \mathbf{R}^{m \times o-m}$ and $\mathbf{B}_d \in \mathbf{R}^{m
\times m}$. With this in mind, equation (\ref{eq:constraint_Bu0}) can be
written as
\begin{align}
\mathbf{B}_i \mathbf{u}_i + \mathbf{B}_d \mathbf{u}_d + \mathbf{f}_{vt}(\mathbf{q}, t) &= \mathbf{0} \\
\implies \mathbf{u}_d &= -\mathbf{B}_d^{-1} \mathbf{B}_i \mathbf{u}_i - \mathbf{f}_{vt}(\mathbf{q}, t)
\end{align}
For convenience, we define
\begin{align}
\label{eq:constraint_A}
\mathbf{A} \triangleq - \mathbf{B}_{d}^{-1} \mathbf{B}_i
\end{align}
and note that $\mathbf{A}\in\mathbf{R}^{m \times (o - m)}$. To constrain the
generalized active and inertia forces, we partitioned them in a similar
manner
\begin{align}
\label{eq:F_Fstar_rewrite}
\mathbf{F} = \begin{bmatrix} \mathbf{F}_i \\ \mathbf{F}_d \end{bmatrix} \quad
\quad \mathbf{F}^* = \begin{bmatrix} \mathbf{F}^*_i \\ \mathbf{F}^*_d
\end{bmatrix}
\end{align}
allowing the nonholonomic generalized forces to be written:
\begin{align}
\label{eq:F_Fstar_nonholonomic}
\tilde{\mathbf{F}} = \mathbf{F}_i + \mathbf{A}^T \mathbf{F}_d \quad \quad
\tilde{\mathbf{F}}^* = \mathbf{F}_i^* + \mathbf{A}^T \mathbf{F}_d^*
\end{align}
which allows Kane's nonholonomic dynamical equations to be written as
\begin{align}
\label{eq:kanes_eq_nonholonomic}
\tilde{\mathbf{F}} + \tilde{\mathbf{F}}^* = \mathbf{0}
\end{align}
This process of constraining the equations of motion reduces the number of
equations from $o$ to $o-m$. However, eqns. (\ref{eq:kanes_eq_nonholonomic}) still
contain $o$ unknowns ($\dot{\mathbf{u}}_1,...,\dot{\mathbf{u}}_o$). By
augmenting (\ref{eq:kanes_eq_nonholonomic}) with the acceleration constraint
equations (\ref{eq:acceleration_constraints}), we obtain $o$ equations in
$o$ unknowns. Note that equations (\ref{eq:kanes_eq_nonholonomic}) and
(\ref{eq:acceleration_constraints}) are linear in the $\dot{u}_i$'s. The
augmented system of equations is
\begin{align}
\label{eq:F_Fstar_f_a}
\begin{bmatrix}
\tilde{\mathbf{F}} + \tilde{\mathbf{F}}^* \\
\mathbf{f}_a (\mathbf{q}, \dot{\mathbf{q}}, \mathbf{u}, \dot{\mathbf{u}}, t)
\end{bmatrix} = \mathbf{0}
\end{align}

@@

Not all $n$ generalized coordinates are independent due to the $l$
configuration level constraints (e.g., holonomic constraints resulting from
closed kinematic loops). Similarly, not all $o$ generalized speeds are
independent due to the $m$ velocity level constraints.  Velocity level
constraints result from time differentiated configuration constraints or
nonholonomic constraints (e.g., no-slip rolling).  The $m$ acceleration level
constraints are assumed to be obtained by time differentiating the $m$ velocity
level constraints.  Just as Kane's method does not require $n = o$, neither
does our linearization procedure. The $s$ exogenous inputs include control
system inputs and/or external disturbance forces or torques applied to the
system.

The kinematic differential equations are separated into the quantities
$\mathbf{f}_0$ and $\mathbf{f}_1$; terms depending upon $\dot{\mathbf{q}}$ are
placed into $\mathbf{f}_0$, all other terms are placed into $\mathbf{f}_1$.
Similarly, the dynamic differential equations are separated into the
quantities $\mathbf{f}_2$ and $\mathbf{f}_3$; terms depending upon
$\dot{\mathbf{u}}$ are placed into $\mathbf{f}_2$, all other terms are placed
into $\mathbf{f}_3$. The motivation for this organizational scheme will be made
apparent in Section \ref{sec:derivations}.

\section{Dynamics Example}
\label{sec:example}
To illustrate Kane's method in the presence of configuration and velocity
constraints, we consider a thin disk $C$ rolling without slip on a horizontal
plane. The dextral unit vectors $\hat{c}_x, \hat{c}_y, \hat{c}_z$ are fixed to $C$
with $\hat{c}_y$ perpendicular to the disk plane. Let $r$ and $m$ be the
disk radius and mass, and assume a thin disk mass distribution so that
$I_{xx}=I_{zz}=mr^2/4$ and $I_{yy} = mr^2/2$. Take the inertial (Newtonian)
frame $N$ to be similarly equipped with dextral unit vectors $\hat{n}_x,
\hat{n}_y, \hat{n}_z$, with $\hat{n}_z$ perpendicular to the ground plane
(downwards) and aligned with the local gravitational field. To orient the disk,
first align $C$ with $N$, then apply successive body fixed ZXY rotations:
yaw $q_1$, roll $q_2$, and spin $q_3$.  Denote by $A$ and $B$ the two frames
which persist after the first two intermediate rotations in the sequence; we
refer to these as the instantaneous yaw and lean frames, respectively. The unit
vector, $\hat{b}_z$ fixed in the $B$ frame, is parallel to the disk plane and
aligned with the position vector from the disk center $C^*$ to the lowest point
of the disk $P^*$. Figure \ref{fig:rolling_disc} shows how the generalized
coordinates are used to configure the disc.
